Conditional selectivity for triplestores

### 2012-09-17 conditional/trunk/selectivity-basic.tex
headline
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@368 43ef1190-90fd-4dd7-a769-d0121f02b80e
define \sel macro
Comment out blatantly flawed stuff


git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@369 43ef1190-90fd-4dd7-a769-d0121f02b80e

\newcommand{\rdfterm}[1]{\texttt{#1}}
\newcommand{\todo}[1]{\ensuremath{^{\textrm{\tiny{TODO}\normalsize}}}\footnote{\textbf{TODO:}~#1}}

@@

Analogously\todo{anybody good at arguing this mathematically?}, the
conditional selectivity for a triple pattern $T$ can be written as 
\begin{equation}\label{eq:sel}
sel(T) = sel(S \cap P \cap O) = sel(O | S \cap P) \cdot sel(P | S)
\cdot sel(S) . 
\end{equation}
We note that intersection is commutative, thus similar
relations can be shown for each permutation of terms.

@@

A very simple example is to find the selectivity for the triple pattern
\begin{verbatim}
?person a foaf:Person .
\end{verbatim}
We can easily see that the correct answer is $sel(T) = 2/7$ since two
out of the 7 triples match the pattern. To see how Eq.~\ref{eq:sel} is
used, we note that $sel(S) = 1$ since it is a variable. $sel(P | S) =
2/7$ since the predicate matches the two occurences of
\rdfterm{rdf:type} and the subject variable doesn't influence
this. Finally, $sel(O | S \cap P) = 1$ since the only two occurences
of \rdfterm{foaf:Person} occurs in the same triples as
\rdfterm{rdf:type}. Thus, we arrive at the expected result.

In a slightly harder example:
\begin{verbatim}
?person foaf:firstName "Bob" .
\end{verbatim}
We still have $sel(S) = 1$ as above, and there are still two
predicates that match, so $sel(P | S) = 2/7$, but only half of those
match the literal, so $sel(O | S \cap P) = 1/2$, giving us the
expected $sel(T) = 1/7$.

@@

@@

\begin{equation}\label{eq:selsubjun}
sel(T_{?~p~o}) = sel(O | P) \cdot sel(P) . 
\end{equation}

@@

Since property partitions detail statistics given a certain predicate,
they can be used in the conditional case. We can therefore write:

\begin{equation}\label{eq:selsubjunp}
sel(P = p) = \frac{\rdfterm{void:propertyPartition/void:triples}}{|d|}
\end{equation}
and
\begin{equation}\label{eq:selsubjunop}
sel(O = o|P = p) =
\frac{\rdfterm{void:propertyPartition/void:distinctObjects}}{\rdfterm{void:propertyPartition/void:triples}} .
\end{equation}

Thus,
\begin{equation}\label{eq:selsubjunop}
sel(T_{?~p~o}) =
\frac{\rdfterm{void:propertyPartition/void:distinctObjects}}{|d|} .
\end{equation}\todo{wtf}

@@

Then, 
\begin{equation}
sel(S = s_i) = \frac{cm(s_i)}{i} ,
\end{equation}
\begin{equation}
sel(P = p_{n_i}^i|S = s_i) = \frac{cm(p_{n_i}^i)}{n_i}
\end{equation}
and
\begin{equation}
sel(O = o_{k_{i,n_i}}^{i,n_i} |S = s_i \cap P = p_{n_i}^i) =
\frac{cm(o_{k_{i,n_i}}^{i,n_i})}{k_{i,n_i}} .
\end{equation}
\todo{surely needs more work}

>>>>

\newcommand{\rdfterm}[1]{\texttt{#1}}
\newcommand{\todo}[1]{\ensuremath{^{\textrm{\tiny{TODO}\normalsize}}}\footnote{\textbf{TODO:}~#1}}
\newcommand{\sel}[1]{\ensuremath{sel\left(#1\right)}}

@@

Analogously\todo{anybody good at arguing this mathematically?}, the
conditional selectivity for a triple pattern $T$ can be written as 
\begin{equation}\label{eq:sel}
\sel{T} = \sel{S \cap P \cap O} = \sel{O | S \cap P} \cdot \sel{P | S}
\cdot \sel{S} . 
\end{equation}
We note that intersection is commutative, thus similar
relations can be shown for each permutation of terms.

@@

A very simple example is to find the selectivity for the triple pattern
\begin{verbatim}
?person a foaf:Person .
\end{verbatim}
We can easily see that the correct answer is $\sel{T} = 2/7$ since two
out of the 7 triples match the pattern. To see how Eq.~\ref{eq:sel} is
used, we note that $\sel{S} = 1$ since it is a variable. $\sel{P | S} =
2/7$ since the predicate matches the two occurences of
\rdfterm{rdf:type} and the subject variable doesn't influence
this. Finally, $\sel{O | S \cap P} = 1$ since the only two occurences
of \rdfterm{foaf:Person} occurs in the same triples as
\rdfterm{rdf:type}. Thus, we arrive at the expected result.

In a slightly harder example:
\begin{verbatim}
?person foaf:firstName "Bob" .
\end{verbatim}
We still have $\sel{S} = 1$ as above, and there are still two
predicates that match, so $\sel{P | S} = 2/7$, but only half of those
match the literal, so $\sel{O | S \cap P} = 1/2$, giving us the
expected $\sel{T} = 1/7$.

@@

\subsection{Terms that can be computed}

\begin{equation}\label{eq:selpp}
\sel{P = p} = 
\end{equation}

\subsection{Triple pattern cases}

@@

% \begin{equation}\label{eq:selsubjun}
% \sel{T_{\{?~p~o\}}) = \sel{O | P) \cdot \sel{P) . 
% \end{equation}

@@

% Since property partitions detail statistics given a certain predicate,
% they can be used in the conditional case. We can therefore write:

% \begin{equation}\label{eq:selsubjunp}
% \sel{P = p) = \frac{\rdfterm{void:propertyPartition/void:triples}}{|d|}
% \end{equation}
% and
% \begin{equation}\label{eq:selsubjunop}
% \sel{O = o|P = p) =
% \frac{\rdfterm{void:propertyPartition/void:distinctObjects}}{\rdfterm{void:propertyPartition/void:triples}} .
% \end{equation}
% Thus,
% \begin{equation}\label{eq:selsubjunres}
% \sel{T_{\{?~p~o\}}) =
% \frac{\rdfterm{void:propertyPartition/void:distinctObjects}}{|d|} ,
% \end{equation}\todo{WRONG!}
% where the notation is inspired by SPARQL 1.1 Property Paths,
% for example the relevant part of a VoID description for the second
% example in section~\ref{sec:example} would be:
% \begin{verbatim}
% </dataset> a void:Dataset ;
%            void:propertyPartition [
%                 void:property foaf:firstName;
%                 void:distinctObjects 2 .
%            ]
% \end{verbatim}

% there exists a \rdfterm{void:propertyPartition} predicate that
% has an object which is again linked to a dataset subset containing the
% property $p$ as predicate of

@@

Then, 
\begin{equation}
\sel{S = s_i} = \frac{cm(s_i)}{i} ,
\end{equation}
\begin{equation}
\sel{P = p_{n_i}^i|S = s_i} = \frac{cm(p_{n_i}^i)}{n_i}
\end{equation}
and
\begin{equation}
\sel{O = o_{k_{i,n_i}}^{i,n_i} |S = s_i \cap P = p_{n_i}^i} =
\frac{cm(o_{k_{i,n_i}}^{i,n_i})}{k_{i,n_i}} .
\end{equation}
\todo{surely needs more work}
